---
title: "R Notebook"
output:
  html_notebook: default
  pdf_document: default
---

This exercise prepares data from a study that was conducted in the 3 districts that make up Kigali City. 
In each of the districts, a list of households in ubudehe category 1 was collected.
For each household, additional information was collected on the household head.
There are 2 datasets for each district:
• a main dataset, and
• a roster dataset.
The main datasets contain information on:
i. where the households on which ubudehe data was collected are located,
ii. the local authority in charge of ubudehe data, and
iii. the number of households in ubudehe category 1.
The roster dataset contains information on:
i. the household head in each of the households listed.


```{r}
#import R packages
library(tidyverse)
library(labelled)
library(openxlsx)
library(dbplyr)
```

```{r}
#create dataframes to store data
main <- read.csv("C:\\Users\\user\\Downloads\\laterite_dq analyst_technical assessment_data cleaning_sept 2023\\main_dataset.csv", header=TRUE, stringsAsFactors=FALSE)
roster <- read.csv("C:\\Users\\user\\Downloads\\laterite_dq analyst_technical assessment_data cleaning_sept 2023\\roster_dataset.csv", header=TRUE, stringsAsFactors=FALSE)

```



```{r}
#label all columns in the main dataset
var_label(main) <- c('province_id'='Province code',
'province_name'='Province of residence',
'district_id'='District code',
'district_name'='District of residence',
'sector_id'='Sector code',
'sector_name'='Sector of residence',
'cell_id'='Cell code',
'cell_name'='Cell of residence',
'village_id'='Village code',
'village_name'='Village of residence',
'list_available'='If the ubudehe list is available',
'list_source'='From whom the ubudehe list was obtained',
'ubudehe_contact'='Ubudehe list contact person',
'contact_role'='Role of the ubudehe list contact person',
'contact_phone'='Phone number of ubudehe list contact person',
'num_households'='Number of households in ubudehe',
'gpslatitude'='Latitutde of where list was obtained',
'gpslongitude'='Longitude of where list was obtained',
'gpsaltitude'='Altitude of where list was obtained',
'gpsaccuracy'='Accuracy of the GPS coordinates obtained',
'parent_key'='Unique ID')
```



```{r}
#label all columns in the roster dataset
var_label(roster) <- c('hh_head_position'='Position of household head',
'name_hh_head'='Name of the household',
'nid_hh_head'='National ID number of the household head',
'phone_hh_head'='Phone number of household head',
'parent_key'='Unique ID')
```



```{r}
#assign labels to values of the categorical variables 
val_labels(main$list_available) <- c(Yes = 1, No = 0)
val_labels(main$list_source) <- c('Sector Executive Officer'=1,
                                         'Cell Executive Officer'=2,
                                         'Village Leader'=3)
val_labels(main$contact_role) <- c('Village leader'=1,
                                         'Village in-charge of security'=2,
                                         'Village Community Health Worker'=3)
```


```{r}
#assign values of the variable province_name as the labels of the values of the variable province_id
provinces <- deframe(unique(main[c("province_name","province_id")]))
val_labels(main$province_id) <- provinces
```



```{r}
#assign values of the variable district_name as the labels of the values of the variable district_id
districts <- deframe(unique(main[c("district_name","district_id")]))
val_labels(main$district_id) <- districts
```



```{r}
#assign values of the variable sector_name as the labels of the values of the variable sector_id
sectors <- deframe(unique(main[c("sector_name","sector_id")]))
val_labels(main$sector_id) <- sectors
```


```{r}
#assign values of the variable cell_name as the labels of the values of the variable cell_id
cells <- deframe(unique(main[c("cell_name","cell_id")]))
val_labels(main$cell_id) <- cells
```


```{r}
#assign values of the variable village_name as the labels of the values of the variable village_id
villages <- deframe(unique(main[c("village_name","village_id")]))
val_labels(main$village_id) <- villages
```


```{r}
#create new location variables as factors so that we can display the value labels in place of the numerical values
main$province <- as_factor(main$province_id,levels ="labels")
main$district <- as_factor(main$district_id,levels ="labels")
main$sector <- as_factor(main$sector_id,levels ="labels")
main$cell <- as_factor(main$cell_id,levels ="labels")
main$village <- as_factor(main$village_id,levels ="labels")
```


```{r}
#merge the main and roster datasets by adding a new column to the main dataset which contains combined roster data of all households located in each village
merged_agneswambui_20231231 <- main #copy main dataset
hh_details = list() #create empty list to store roster data
key <- main$parent_key #
index <- 1
for (k in key){ #loop through all villages and combine their roster data
  hh_details[[index]] <- roster[roster$parent_key == k,]
  index <- index+1
}
 
merged_agneswambui_20231231$hh_details <- hh_details #create new dataset
```


```{r}
#combine the latitude and longitude variables to create a gps coordinates variable
clean_agneswambui_20231231 <- merged_agneswambui_20231231 #copy merged data
clean_agneswambui_20231231$gps <- paste(merged_agneswambui_20231231$gpslatitude,merged_agneswambui_20231231$gpslongitude)
```


```{r}
#reorder columns
clean_agneswambui_20231231 <- clean_agneswambui_20231231[,c('province_id','province_name','district_id','district_name','sector_id','sector_name','cell_id','cell_name','village_id','village_name','list_available','list_source','ubudehe_contact','contact_role','contact_phone','num_households','hh_details','gps','parent_key')]

```


```{r}
#assign labels to new variables
var_label(clean_agneswambui_20231231$gps) <- 'GPS coordinates of where the list was obtained'
var_label(clean_agneswambui_20231231$hh_details) <- 'Personal information of heads of households'
```


```{r}
#check that all variables have been labelled correctly
var_label(clean_agneswambui_20231231)
```

```{r}
#generate an excel file of roster data for each district and put data for different sectors in separate worksheets  
districts <- unique(clean_agneswambui_20231231$district_name) #get unique list of districts
for (d in districts){ #loop through each district
  sector_data <- list()
  index_districts <- 1 
  data <- clean_agneswambui_20231231[clean_agneswambui_20231231$district_name == d,]
  sectors <- unique(data$sector_name) #get unique list of sectors
  
  for (s in sectors){ #loop through each sector in the district
    hh_details <- data[data$sector_name == s,]$hh_details
    index_districts <- index_districts+1
    df_list = list()
    
    for (f in 1:length(hh_details)){ #loop through each village in the sector
      df_list[[f]] <- hh_details[f]
      
    } 
    df_combine <- bind_rows(df_list) #combine roster data for each sector
    sector_data[[index_districts]] <- df_combine 
  }
  path <- paste0("C:\\Users\\user\\Downloads\\laterite_dq analyst_technical assessment_data cleaning_sept 2023\\",d,".xlsx")
  write.xlsx(sector_data, file = path) #create excel file to save data for each sector on separate worksheets and repeat for all districts
}
```





